`r params$project_title`

### 2020-10-15 reports/rmarkdown/01_data_pre_processing.Rmd
updates to export uncorrected data CVL2

As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits the data and learns the structure of the data.

```{r, echo = FALSE}
batch2_all <- t(batch2[nonzeros,])
batch2_all <- batch2_all/max(batch2_all)

@@

datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)

write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )

>>>>

As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits and learns the structure of the data.

```{r, echo = FALSE,fig.height=5,fig.width=5}
datasets[["ASV_CVL_V2_raw"]] <- cbind( datasets[["ASV_CVL_V2_B2.csv"]] , datasets[["ASV_CVL_V2_B1.csv"]], datasets[["ASV_CVL_V2_C.csv"]][,! colnames(datasets[["ASV_CVL_V2_C.csv"]]) %in% common_samples] )

batch2_all <- t(batch2[nonzeros,])
batch2_all <- batch2_all/max(batch2_all)

@@

datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2_raw"]] <- round(t( log2( t(datasets[["ASV_CVL_V2_raw"]]) / colSums(datasets[["ASV_CVL_V2_raw"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)

write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )

batches_CVL2 <- ifelse( colnames(datasets[["ASV_CVL_V2_raw"]]) %in% sub(".*_","",colnames(batch1)),"boston","CTMR" )
batches_CVL2 <- t(t(setNames(batches_CVL2 , colnames(datasets[["ASV_CVL_V2_raw"]]))))
write.csv( batches_CVL2 , paste0(PATH,"../results/batches_CVL2.csv"),row.names = T )
```

### 2020-10-15 reports/rmarkdown/02_data_quality_control.Rmd
updates to export uncorrected data CVL2

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

@@

# Load datasets
dataset_names <- c("Tissue_RNAseq_V3_normalized",
                     "ASV_tissue_V3_normalized_batch_corrected",
                     "ASV_CVL_V3_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_batch_corrected")

@@

sum((colSums(counts == 0)>=3)*1)
```

sum((colSums(counts == 0)>=3)*1)
```

>>>>

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)

#Graph construction  
fct <- .5 # FC threshold for differential expression
pvt <- 0.01 # Pvalue threshold for differential expression
min_pct <- .1 # minimun level of detected bacteria in each sample group
```

@@

# Load datasets
dataset_names <- c("Tissue_RNAseq_V3_normalized",
                     "ASV_tissue_V3_normalized_batch_corrected",
                     "ASV_CVL_V3_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_NOT_batch_corrected")

@@

sum((colSums(counts == 0)>=3)*1)
colnames(counts)[(colSums(counts == 0)>=3)]

metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
```

# Comparing CVL2 samples before and after batch correction

```{r, echo = FALSE}
mypar(1,2,mar=c(2,15,1,1))

for( dataset in c("ASV_CVL_V2_normalized_NOT_batch_corrected",
            "ASV_CVL_V2_normalized_batch_corrected")){

  # Computing differential expression across batches
  all_microbiome <- cbind(datasets_all_samples[[dataset]])
  datasets <- read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,2]

  all_microbiome <- all_microbiome[,as.character(read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,1])]

  NN <- min(table(datasets))

  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
  for(i in levels(datasets)){
    for(j in rownames(all_microbiome) ){

      set.seed(1)
      a <- c(sample(all_microbiome[j,datasets == i],NN))
      set.seed(1)
      b <- sample(all_microbiome[j,datasets != i],NN)

      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

      temp <- wilcox.test(x=a, y=b,exact = F)

      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                            c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
    }
  }

  res <- res[-1,]
  res$pvalue <- as.numeric(res$pvalue)
  res$perc.1 <- as.numeric(res$perc.1)
  res$perc.2 <- as.numeric(res$perc.2)
  res$perc.diff <- res$perc.1 - res$perc.2

  res$fc <- as.numeric(res$fc)
  res$FDR <- p.adjust(res$pvalue)
  res <- res[order(res$pvalue),]
  res <- res[res$fc > fct,]
  res <- res[abs(res$pvalue) < 0.05,]
  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
  dim(res)

  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets, main = dataset,show_grid = T,cex.main=1,font.main=1,
            cex.row = .8,cex.col = .8,srt = 0)
}
```

```{r, echo = FALSE}
# Computing differential expression across datasets

all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    set.seed(1)
    a <- c(sample(all_microbiome[j,datasets == i],NN))
    set.seed(1)
    b <- sample(all_microbiome[j,datasets != i],NN)

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(mar=c(2,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)

# Correlation CVL3 and RNAseq

```{r, echo = FALSE}

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
CVL3 <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized","ASV_CVL_V3_normalized_batch_corrected"),] > 0)==2 ]

#filter TRX dataset
TRX <- TRX [ , samples_TRX_and_CVL3 ]
TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:3000] )
plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
TRX <- TRX[ top_vars_TRX , ]
dim(TRX)

#filter CVL3 dataset
CVL3 <- CVL3 [ , samples_TRX_and_CVL3 ]
CVL3 <- CVL3 [ rowSums(CVL3>0)>= 2 , ]
top_vars_CVL3 <- names(sort(apply(CVL3,1,var),decreasing = T)[1:100] )
dim(CVL3)

mypar()
cors <- cor( t(rbind(TRX) ) ,  t(rbind(CVL3) ) )
hist(cors,xlim=c(-1,1),breaks = 100)
dim(cors)

library(fgsea)
bacteria_use <- colnames(cors)
gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
enrichments <- lapply(bacteria_use,function(x){
  res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=1000)
  return(res)
})
names(enrichments) <- bacteria_use

pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
pvalues <- as.data.frame(pvalues)
pvalues <- -log10( pvalues )
top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
mypar(mar=c(4,15,2,2))
boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )

pvalues[ pvalues < -log10(0.05)] <- 0
sort(rowSums(pvalues>0),T)
sort(colSums(pvalues>0),T)

NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
NESes <- t(as.data.frame(NESes))
# 
# o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
# o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
o_kegg <- hclust( dist( t(NESes) ),"ward.D2")$order
o_bacs <- hclust( dist( NESes ),"ward.D2")$order
NESes <- NESes[o_bacs,o_kegg]

mypar(mar=c(15,2,2,15))
image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-4,4,length.out = 92),
       axes=F,border=NA,main="KEGG pathways",xlab="",ylab="bacterias",line=.4,cex.main=1,font.main=1)
text(  par("usr")[c(4)]+.02 , seq(1,0,length.out = nrow(NESes)),
      labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.3)
text( seq(0,1,length.out = nrow(NESes)) , par("usr")[c(1)]-.02,
      labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.3)

eee <- data.frame(x=c(sapply(colnames(NESes),function(x)rep(x,nrow(NESes)))),
                  y=rep(rownames(NESes),ncol(NESes)),
                  w=c(NESes),
                  p=c(t(pvalues)),
                  x1=0,
                  y1=c(sapply(ncol(NESes):1,function(x)rep(x,nrow(NESes))))/ncol(NESes) ,
                  x2=1,
                  y2=rep(nrow(NESes):1,ncol(NESes))/nrow(NESes) )

eee$w2 <- eee$w / max(eee$w)
eee <- eee[eee$w > 1.5,]
eee <- eee[eee$p > 2,]

eee$y1 <- rank(eee$y1) / length(eee$y1)
eee$y2 <- rank(eee$y2) / length(eee$y2)

plot(c(0,1),c(0,1),type="n")
for( i in 1:nrow(eee)){
  lines(c(eee$x1[i],eee$x2[i]),c(eee$y1[i],eee$y2[i]),
        col=colorRampPalette(c("grey95","grey","firebrick"))(99)[ eee$w2 * 98 + 1 ] )
}

```{r, echo = FALSE}

g <- graph_from_incidence_matrix( 1-cors, directed = F,weighted = T  )

l <- layout_with_fr(g)
plot( g , vertex.label.cex=0.000001 , layout=l,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=1,vertex.color=ifelse(V(g)$name %in% colnames(CVL3),"red","blue") )

### 2020-10-15 reports/rmarkdown/report.Rmd
updates to export uncorrected data CVL2

#Graph construction  
k <- 10 # k for KNN
P <- 0.4 # prunning threshold of SNN

@@

# Load datasets
dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_batch_corrected.csv")

@@

# Create empty matrix for the full dataset
microbiome <- as.matrix(datasets[[3]])
microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]

>>>>

#Graph construction  
k <- 10 # k for KNN
P <- 0.35 # prunning threshold of SNN

@@

# Load datasets
dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_NOT_batch_corrected.csv")

@@

# Create empty matrix for the full dataset
microbiome <- as.matrix(datasets[[4]])
microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]
